---
title: "Introduction to Functions in R"
author: "William Murrah"
date: "11/16/2014"
output: ioslides_presentation
bibliography: bibfile.bib
---

## Anatomy of an R Function

```{r eval=FALSE}

FunctionName <- function(arguments){ 
  statements that do stuff,
  these are R code
}
  
```

see [@gcode] for conventions on how to name functions. Function names should be verbs that describe what the function does. They should be capitalized and use *camel case*.

## A Basic Example: Creating an `Add` Function

>
```{r}
Add <- function(x, y) {
  x + y
}
```

>
```{r}
Add(3, 4)
```


## Create a `Division` function

```{r}
Division <- function(numerator, denominator=2) {
  numerator/denominator
}
```

## `Division` Examples

```{r}
Division(1, 2)
Division(1)
Division(2)

```

## `Division` Examples
```{r}
Division(2, 4)
Division(numerator=2, denominator=4)
Division(denominator=4, numerator=2)
Division(4, 2)
```

## Creating the `Average` Function

```{r }
Average <- function(x) {
sum(x)/length(x)
}
```


```{r}
vec <- 0:100
Average(vec)
```

## `for` Loops

```{r}
x  <- c(1, 4, NA, 1, 1)

```

## `for` Loops

```{r}
x  <- c(1, 4, 3, 1, 1)
N <- 0

```

## `for` Loops

```{r}
x  <- c(1, 4, 3, 1, 1)
N <- 0
  for (i in 1:5){
    N  <- N + x[i]
    }

```



## `for` Loops

```{r}
x  <- c(1, 4, 3, 1, 1)
N <- 0
  for (i in 1:5){
    N  <- N + x[i]
    }
N
```


## `Average` Revisited | with `for` loop

```{r}
Average <- function(x) {
  N <- 0
  for (i in 1:length(x)){
    N  <- N + x[i]
    }
  avg <- N/length(x)
  return(avg)
}

```

## `Average` Revisited | with `for` loop

```{r}
Average <- function(x) {
  N <- 0
  for (i in 1:length(x)){
    N  <- N + x[i]
    }
  avg <- N/length(x)
  return(avg)
}

```

## `Average` Function Compared to `base::mean` Function

```{r}
x <- 0:100
Average(x)
mean(x)
```



## `Average` returing list
```{r}
Average <- function(x) {
  N <- 0
  for (i in 1:length(x)){
    N  <- N + x[i]
    }
  avg <- N/length(x)
  
  return(list(sum.x=N, average=avg))
}

Average(x)
```

## References