---
title: "R Markdown Tutorial"
author: "William Murrah"
date: "02/13/2015"
output: pdf_document
csl: apa.csl
bibliography: bibfile.bib
---

This tutorial was developed to introduce DARA Team members to R Markdown. It focuses on topics relevant to what we do in the FOCAL lab, emphasizing the skills we will need to generate reports. I will not go over the basics of the `Markdown` markup language. That is covered well in the help files, which can be accessed by clikcin the '?' in the menu of an R Markdown document. 

```{r opts, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(data.table)
library(tables)
opts_knit$set(root.dir='../../')
opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)              
```


```{r data}
# getwd()
mtcars <- read.csv(file ="data/mtcars.csv", header = TRUE)
```


```{r ,results='asis'}
latex(
tabular(Factor(cyl) ~ Factor(gear), data=mtcars)
)
```

I read this really cool article the other day [@kim2014psychometric].

@cameron2012fine said some cool stuff.

```{r scatterplot}
plot(mpg ~ wt, data=mtcars, pch=20, col="blue")
```


# References