---
title: "Intraclass correlations"
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
---

## Set up 
```{r, message=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, comment=NULL)
library(lme4)
library(psychometric)
library(psych)
library(texreg)
library(ztable)
library(xtable)
library(mosaic)
options(ztable.type="html")
options(digits = 2)
```


## Extreme 1: Very Low Within Group Variance, High Between Group Variance

## Data

Create a data set with very low within group variance and high between group variance.
```{r, results='asis'}
dat <- data.frame(Y = c(seq(0,1, length.out = 10), 
                        seq(4.5, 5.5, length.out =  10), 
                        seq(9,10, length.out =  10)), 
                  Group = factor(rep(1:3, each=10)))

# Print data frame.
z <- ztable(dat, align = "ccc")
z <- addrgroup(z, 
               rgroup = c("Group 1", "Group 2", "Group 3"), 
               n.rgroup = c(10, 10, 10),
               cspan.rgroup=1)
z
```

## Boxplot of data
```{r}
boxplot(Y ~ Group, 
        data = dat, 
        xlab="Group", 
        ylab="Y", 
        main= "Low within group variance, high between group variance")
```

## Descriptive statistics of data
```{r}
mean(dat$Y)
var(dat$Y)
sd(dat$Y)
aggregate(Y ~ Group, 
          data = dat, 
          FUN = function(x) c(Mean=mean(x), 
                              VAR=var(x), 
                              SD=sd(x)))
```

## Variance of the group means
```{r}
var(c(.5, 5, 9.5))
```

## Intraclass correlation
```{r}
ICC1.lme(Y, Group, dat)
```



## Comparison of models of clustered data
```{r, results='asis'}
# Linear model ignoring clustered nature of data.
lm.mod <- lm(Y ~ 1, 
             data = dat)

# Linear model with fixed effects for group.
fe.mod <- lm(Y ~ Group, 
             data = dat)
# Hierarchical model.
ml.mod <- lmer(Y ~ 1 + (1 | Group), 
               data = dat)

# Compare models
htmlreg(l=list(lm.mod, fe.mod, ml.mod), 
        custom.model.names = c("Linear", "Fixed Effects", "Hierarchical"),
        caption = "Comparing models of clustered data" ,
        caption.above=TRUE)
```


* * *
# Extreme 2: High Within Group Variance, Low Between Group Variance
```{r, results='asis'}
dat2 <- data.frame(Y = c(seq(0,10, length.out = 10) - .1, 
                         seq(0,10, length.out = 10), 
                         seq(0,10, length.out = 10) + .1), 
                  Group = factor(rep(1:3, each=10)))

# Print data frame.
z2 <- ztable(dat2, align = "ccc")
z2 <- addrgroup(z2, 
                rgroup = c("Group 1", "Group 2", "Group 3"), 
                n.rgroup = c(10, 10, 10),
                cspan.rgroup=1)
z2
```

## Boxplot of data
```{r}
boxplot(Y ~ Group, 
        data = dat2, 
        xlab="Group", 
        ylab="Y", 
        main= "High within group variance, no between group variance")
```

## Descriptive Statistics
```{r}
mean(dat2$Y)
var(dat2$Y)
sd(dat2$Y)
aggregate(Y ~ Group, 
          data = dat2, 
          FUN = function(x) c(Mean=mean(x), VAR=var(x), SD=sd(x)))

```

## Variance of the group means
```{r}
var(c(4.9, 5.0, 5.1))
```


```{r}
ICC1.lme(Y, Group, dat2)
```


```{r, results='asis'}
# Linear model ignoring clustered nature of data.
lm.mod2 <- lm(Y ~ 1, 
              data = dat2)

# Linear model with fixed effects for group.
fe.mod2 <- lm(Y ~ Group, 
              data = dat2)

# Hierarchical model.
ml.mod2 <- lmer(Y ~ 1 + (1 | Group), 
                data = dat2)

htmlreg(l=list(lm.mod2, fe.mod2, ml.mod2), 
        custom.model.names = c("Linear", "Fixed Effects", "Hierarchical"),
        caption = "Comparing models of clustered data" ,
        caption.above=TRUE)
```


```{r}
par(mfrow=c(1,2))
boxplot(Y ~ Group, dat, xlab="Group", ylab="Y", main= "Low within, high between")

boxplot(Y ~ Group, dat2, xlab="Group", ylab="Y", main= "High within no between")

```

## Low within, High between
```{r}
# par(mfrow = c(3,1))
plotreg(list(lm.mod, fe.mod, ml.mod), 
        xlim = c(-3, 10), 
        mfrow = FALSE, 
        custom.model.names = c("Linear", "Fixed Effects", "Hierarchical"))

```

## High within, low between
```{r}
plotreg(list(lm.mod2, fe.mod2, ml.mod2), 
        xlim = c(-3, 10), 
        mfrow =FALSE, 
        custom.model.names = c("Linear", "Fixed Effects", "Hierarchical"))
```

