---
title: "Introduction to Multilevel Modeling"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
header-includes:
- \usepackage{amsmath}
bibliography: references.bib    
---


```{r, include=TRUE, message=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
# Other packages to load:
require(ztable)
require(texreg)
require(lme4)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
require(mosaicData)  
options(ztable.type = "html")
```

# Basic Notation

$$
Y_{ij} = \beta_{0j} + e_{ij}, \tag{1}
$$

$$
\beta_{0j} = \gamma_{00} + u_{0j}, \tag{2}
$$

$$
Y_{ij} = \gamma_{00} + u_{0j} + e_{ij}. \tag{3}
$$ 



# Example of Simpson's Paradox

This example of Simpson's paradox is taken from [Joseph Stevens](http://pages.uoregon.edu/stevensj/workshops/hlmslides.pdf), which is based on data from [@snijders1999multilevel].

### Definition from [wikipedia](https://en.wikipedia.org/wiki/Simpson%27s_paradox):
> Simpson's paradox, or the Yule–Simpson effect, is a paradox in probability and statistics, in which a trend that appears in different groups of data but disappears or reverses when these groups are combined. It is sometimes given the impersonal title reversal paradox or amalgamation paradox  .

```{r , results = 'asis'}
toydat <- data.frame(id = 1:10,
                     group = rep(1:5, each = 2),
                     Y = c(5, 7, 4, 6, 3, 5, 2, 4, 1 ,3),
                     X = c(1, 3, 2, 4, 3, 5, 4, 6 ,5, 7))
ztable(toydat, 
       align = "cccc")
```

## Ignore clustering

```{r , results='asis'}
model1 <- lm(Y ~ X, toydat)
htmlreg(model1, 
        caption = "Ignore clustering", 
        custom.model.names = "ignore")
plot(Y ~ X, 
     toydat, 
     main = "Ignore clustering",
     xlim = c(0,8),
     ylim = c(0, 8))
abline(reg = model1, 
        col = "red")


```

$$
Y_i = 5.333 - .33X + e_i
$$

**Interpretation:** There’s a negative relationship between the predictor ***X*** and the outcome ***Y***, a one unit increase in ***X*** results in .333 lower ***Y***.


### Plot of Residuals
```{r , message=FALSE}
mplot(model1, 
      which = 1, 
      system = "ggplot", 
      ncol = 2)
```

## Aggregate to the Group Level
```{r , results= 'asis'}
gdat <- data.frame(group = 1:5,
                   Y = aggregate(Y ~ group,
                                 toydat, 
                                 FUN = mean)$Y,
                   X = aggregate(X ~ group,
                                  toydat,
                                  FUN = mean)$X)
ztable(gdat,
       align = "ccc")
```

```{r , results='asis'}
model2 <- lm(Y ~ X, gdat)
htmlreg(list(model1, model2),
        custom.model.names = c("ignore", "aggregate"))
plot(Y ~ X, 
     gdat, 
     main = "Aggregate",
     xlim = c(0,8),
     ylim = c(0, 8))
abline(reg = model2, 
        col = "red")
```

$$
Y_j = 8.00 - 1.00X + u_j
$$

**Interpretation:** There’s a negative relationship between the predictor ***X*** and the outcome ***Y***, a one unit increase in ***X*** results in 1.0 lower ***Y***.

## Separate Models within Groups

```{r results ='asis'}
sepmodels <- lmList(Y ~ X |group, data = toydat)
htmlreg(sepmodels[1:5])
```

We could also think about this as a fixed effects model

```{r results = 'asis'}
ctr <- function(x) scale(x, scale = FALSE)
femodel <- lm(Y ~ X + as.factor(group), data = toydat )

htmlreg(femodel)
```


```{r results = 'asis'}
plot(Y ~ X, 
     toydat, 
     main = "Separate Group Models",
     xlim = c(0,8),
     ylim = c(0, 8))
modcols = 1:5
for (i in 1:length(unique(toydat$group))) {
  abline(reg = sepmodels[[i]], col = i)
}

```


## Multilevel Modeling

```{r , results='asis'}
mlmodel <- lmer(Y ~ X + (1 | group), data = toydat)
htmlreg(list(model1, model2, femodel, mlmodel),
        custom.model.names = c("ignore", "aggregate", "femodel", "multilevel"))
```




### Document Information
  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * `mosaic` package version: `r packageVersion("mosaic")`
  * Additional session information
  
```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
