---
title: "CKCS Analyses"
author: "William M. Murrah"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
csl: apa.csl
bibliography: tot2.bib
---

```{r opts, warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir='../../../')
opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
library(tables)
library(texreg)
library(tables)
library(Hmisc)
library(AER)
library(sandwich)
library(memisc)
```


```{r data}
source('data/core1Rmake.R')
source('analyses/core1ITT.R')
source('analyses/tslsModelsR1.R')
# source('analyses/coreTSLScontinuous.R')
# source('analyses/coreTSLScontinuousNA0.R')
source('analyses/tslsAEFP_1.R')
source('analyses/tslsAEFP_AK_1.R')

```

# Descriptives

\begin{table}[h]
\center
```{r , results='asis'}
# latex(
#   tabular(Factor(lottery, "Lottery")  ~ Factor(tot.study,name = "Study School Offer", levelnames = c('Declined', 'Accepted')) + 1, core1)
#   )

latex(
  tabular(Factor(lottery, "Lottery")*(n=1)  + (Heading("Column Total")*1)   ~ Factor(tot.study,name = "Study School Offer", levelnames = c('Declined', 'Accepted')) + (Heading("Row Total")*1), core1)
  )

# latex(
#   tabular(Factor(lottery, "Lottery")*((n=1) + Percent(denom="row"))  + (Heading("Column Total")*1)   ~ Factor(tot.study,name = "Study School Offer", levelnames = c('Declined', 'Accepted')) + (Heading("Row Total")*1), core1)
#   )
```
\caption{Crosstabulation of Lottery Offered and Lottery Accepted}
\end{table}

\begin{table}[h]
\center
```{r , results='asis'}
latex(
  tabular(Factor(lottery, "Lottery") + 1   ~ Factor(tot.study,name = "Offer", levelnames = c('Declined', 'Accepted'))*(school + 1) + (Total=1), core1)
  )
```
\caption{Crosstabulation of Lottery Offered and Lottery Accepted by Study School}
\end{table}

\begin{table}[h]
\center
```{r , results='asis'}
latex(
  tabular(Factor(lottery, "Lottery") * (n=1)   + (Heading("Column Total")*1 ) ~ Factor(tot.currentID, "Study School 1st Grade", c("Not Attended", "Attended")) + (Heading("Row Total")*1), core1)
 )
```
\caption{Crosstabulation of Lottery Offered and Study School Attendance at the end of First Grade}
\end{table}

\begin{table}[h]
\center
```{r , results='asis'}
latex(center = 'centering',
  tabular(Factor(lottery, "Lottery") + 1   ~ Factor(tot.currentID, "Study School 1st Grade", c("Not Attended", "Attended"))*(school + 1) + (Total=1), core1)
  )
```
\caption{Crosstabulation of Lottery Offered and Study School Attendance at the end of First Grade by Study School}
\end{table}



\newpage

# Analyses

The primary models for the impact analyses were estimated using the students who applied to one study Core Knowledge charter school. By focusing on this subsample, we can assume one-sided noncompliance, meaning that there are no students who when not offered a placement in a study Core Knowledge charter school would accept a placement in a study Core knowledge charter school. Treatment assignment ($z$) was defined as a binary variable with '1' indicating that the student was offered a placement on one and only one of the study Core knowledge charter schools, and being '0' otherwise. Treatment receipt ($d$) was defined as a binary variable with '1' indicationg the student accepted a placement in one and only one study Core Knowledge charter schools and being '0' otherwise. The outcome variables ($y$) were the PPVT and Woodcock-Johnson scores described in the Methods section.


The $ITT_y$ is the intent-to-treat effect of treatment assignment ($z$) on the outcome ($y$). 

## $ITT_y:$

$$
y_i = \beta_0 + \beta_1 z_i + \epsilon_i.
$$

The $\beta_1$ coefficient is interpreted as the effect of assignment to a study Core Knowledge charter school.

The $ITT_d$ is the reduced form or the first stage equation: 

## $ITT_d:$

$$
d_i = \gamma_0 + \gamma_1 z_i + \nu_i.
$$

The intercept ($\gamma_0$) of this equation can be interpreted as the proportion of subjects who attended a study Core Knowledge charter school even though they were not offered a placement(e.g. Always-takers). Because we are assuming one-sided noncompliance, this value should be zero. The $\gamma_1$ coefficient estimates the proportion of subjects who attend a study charter school in the event that they are offered a placement in a study charter school, minus the proportion that would attend a study charter school even if they would have been assigned to the control group [@gerber2012field]. Because we are assuming one-sided non-compliance, the $\gamma_1$ can be interpreted as the proportion of subjects treated in the event that they are assigned to the treatment condition, which is the proportion of compliers.  


Conceputally, the $LATE_{y(d(z))}:$ is:

$$
 \delta = \frac{\beta_1}{\gamma_1},
$$ 

the intent-to-treat effect divided by the proportion of compliers.

Computationally, the $LATE_{y}$ is estimated by regressing the outcome on the predicted scores from the $ITT_d$ model as follows:
 
## $LATE_{y}:$

$$
y_i = \delta_0 + \delta_1 \hat{d}_i + \upsilon_i,
$$

where $\hat{\delta_1}$ is:

$$
\hat{d}_i = \gamma_0 + \gamma_1 z_i.
$$

We use two-stage least squares regression to estimate this model.

```{r ,results='asis', fig.pos="H"}
itt.d0 <- lm(tot.study ~ lottery + school, core1)
itt.y  <- lm(scale(ppvt1) ~ lottery + school, core1)
late  <- ivreg(scale(ppvt1) ~ tot.study + school, ~ lottery + school, core1)
# late2 <- ivreg(scale(ppvt1) ~ tot.study + school,  ~ 
#                  acsLottery + amaLottery + 
#                  lcaLottery + nvcLottery + praLottery + school,   
#                core1)
hc.d  <- coeftest(itt.d0, vcovHC(itt.d0))
hc.y  <- coeftest(itt.y, vcovHC(itt.y))
hc.late <- coeftest(late, vcovHC(late))
# hc.late2 <- coeftest(late2, vcovHC(late2))
# texreg(list(hc.d, hc.y, hc.late), custom.model.names = c('$ITT_d$', '$ITT_y$', '$LATE_{y(d(z))}$'), caption="PPVT", custom.coef.names = c("(Intercept", "$z$", "Academy", "American", "Lincoln", "New Vision", "Platte River", "K full", "$d$"),
#        reorder.coef= c(1,2,9,3,4,5,6,7,8))
# Without Robust SE.
texreg(list(itt.y, itt.d0, late), custom.model.names = c('$ITT_y$', '$ITT_d$', '$LATE_{y}$'), caption="PPVT", custom.coef.names = c("(Intercept)", "$z$", "Academy", "American", "Lincoln", "New Vision", "Platte River", "$d$"),
       reorder.coef= c(1,2,8,3,4,5,6,7), float.pos='h')

#  xtabs(~lottery + tot.study + school, core1)
```


```{r , results='asis' }

texreg(ckcs.itt, caption = "Intent-to-treat analysis")
```



```{r , results='asis' }
itt.d <- lm(tot.study ~ lottery + school, core1)
tot.cur <- core
itt.d2 <- lm(as.numeric(tot.currentID)-1 ~ lottery + school, core1)
texreg(list(itt.d, itt.d2), caption = "Reduced Form Equation")
```



```{r , results='asis' }
texreg(ckcs.tot2, caption ="Instrumental Variables estimation of Local Average Treatment Effect among the compliers")
```


```{r , results='asis' }
texreg(ckcs.akitt, caption = "Intent-to-treat analysis")
```

```{r , results='asis' }
texreg(ckcs.aktot2, caption ="Instrumental Variables estimation of Local Average Treatment Effect among the compliers for Academic Knowledge subscales")
```
\newpage
```{r2 ,results='asis', fig.pos="H"}
# itt.d0 <- lm(tot.study ~ lotteryOffer + school, core1)
# itt.y  <- lm(scale(ppvt1) ~ lotteryOffer + school, core1)
# late  <- ivreg(scale(ppvt1) ~ tot.study + school, ~ lotteryOffer + school, core1)
# 
# hc.d  <- coeftest(itt.d0, vcovHC(itt.d0))
# hc.y  <- coeftest(itt.y, vcovHC(itt.y))
# hc.late <- coeftest(late, vcovHC(late))
# hc.late2 <- coeftest(late2, vcovHC(late2))
# # texreg(list(hc.d, hc.y, hc.late), custom.model.names = c('$ITT_d$', '$ITT_y$', '$LATE_{y}$'), caption="PPVT", custom.coef.names = c("(Intercept", "$z$", "Academy", "American", "Lincoln", "New Vision", "Platte River", "K full", "$d$"),
# #        reorder.coef= c(1,2,9,3,4,5,6,7,8))
# # Without Robust SE.
# texreg(list(itt.y, itt.d0, late), custom.model.names = c('$ITT_y$', '$ITT_d$', '$LATE_{y}$'), caption="PPVT", float.pos='h',
#        reorder.coef= c(1,2,3,4,5,6,7,13,8,9,10,11,12))
```
\newpage
```{r3 ,results='asis', fig.pos="H"}
# itt.d0 <- lm(tot.study ~ lottery*school, core1)
# itt.y  <- lm(scale(ppvt1) ~ lottery*school, core1)
# late  <- ivreg(scale(ppvt1) ~ tot.study + school, ~ lottery*school, core1)
# 
# hc.d  <- coeftest(itt.d0, vcovHC(itt.d0))
# hc.y  <- coeftest(itt.y, vcovHC(itt.y))
# hc.late <- coeftest(late, vcovHC(late))
# hc.late2 <- coeftest(late2, vcovHC(late2))
# # texreg(list(hc.d, hc.y, hc.late), custom.model.names = c('$ITT_d$', '$ITT_y$', '$LATE_{y}$'), caption="PPVT", custom.coef.names = c("(Intercept)", "$z$", "Academy", "American", "Lincoln", "New Vision", "Platte River", "K full", "$d$"),
# #        reorder.coef= c(1,2,9,3,4,5,6,7,8))
# # Without Robust SE.
# texreg(list(itt.y, itt.d0, late), custom.model.names = c('$ITT_y$', '$ITT_d$', '$LATE_{y}$'), caption="PPVT", float.pos='h',
#        reorder.coef= c(1,2,3,4,5,6,7,13,8,9,10,11,12))
```


\newpage


Table \ref{itt} contains the results from the intent-to-treat (ITT) analysis on students applying to one study charter school. The treatment variable in these analyses is `lotteryOffered`, and indicated whether or not the student was offered attendance in the one charter school applied to in the 2010-2011 school year. Table \ref{tot1} contains the **two-stage least squares** analyses for the same sample of students using the variables created from the parent survey (`tot.ks.st`). Table \ref{tot2} contains the **two-stage least squares** analyses using the dosage variable (`st.attend.2year`) from the enrollment data instead of the treatment receipt variable (`tot.ks.st`) created from the parent survey.

```{r ,results='asis'}
# texreg(ckcs.itt, caption='Intent-to-treat Analyses', label= 'itt')
```

\pagebreak

```{r ,results='asis', }
# texreg(ckcs.tot2, caption='Treatment-on-the-treated Analyses', label = 'tot1')
```


```{r ,results='asis'}
# texreg(ckcs.totNA0, caption='Treatment-on-the-treated Analyses with "continuous" dosage variable, all NAs converted to 0', label='tot2')
```

```{r ,results='asis'}
# texreg(ckcs.tot, caption='Treatment-on-the-treated Analyses with "continuous" dosage variable')
```


\pagebreak

# References
